A Switching Planner for Combined Task and Observation Planning

### 2011-02-07 code/subarchitectures/planner.sa/trunk/src/latex/AAAI11/switching.tex
not sure if its getting shorter, but it's not getting longer

We consider only deterministic-action POMDPs, where all state
uncertainty is expressed in the $(\pp{:init})$ declaration. This
declaration is used by our approach to define a starting state, and
the set of state-assumptions available to sequential planning.
Writing \#\ if the value of a proposition is unspecified, continuing
our example from the previous section, we have the following
assumptions:

@@

\noindent Each assumption corresponds to one distinct {\em
relaxed} visitation of the root term of $(\pp{:init})$. In this
relaxed case, a conjunctive term is visited iff its atomic subterms
are visited, and zero or one of its immediate probabilistic subterms
are visited. For a sequential session, the starting state corresponds
to an assumption (usually unique) with probability $1$. In our
example, the starting state is:

@@

We also annotate DTPDDL model with an assumptive action
$\assumptiveS{i}$ for each element $\prob_i (T_i)$, of each
probabilistic term from $(\pp{:init})$. Here, $\assumptiveS{i}$ can be
executed if no $\assumptiveS{j}$, $j \neq i$, has been executed from
the same probabilistic term, and, either
$(\pp{probabilistic}~..\prob_i~(T_i)..)$ is in the root conjunct, or
it occurs in $T_k$ for some executed $\assumptiveS{k}$. 
%%
We add additional constraints that forbid the scheduling of
assumptions about facts after actions with preconditions, or effects,
that mention those facts. For example, the robot cannot assume it is
plugged into a power source immediately after it unplugs itself.
%%
Executing $\assumptiveS{i}$ in a state $\state$ effects a transition
to a successor state $\state^{T_i}$ with probability $\prob_i$, and
$\state^\bot$ with probability $1 - \prob_i$. Here, $\state^{T_i}$ is
the union of $\state$ with atomic terms from $T_i$. State
$\state^\bot$ is an added sink.

@@

\noindent  Here,  $\reward(\state_i, \action_i)$ is the instantaneous
reward ---i.e., can be negative for action costs, or positive for goal
achievement--- received for executing action $\action_i$ in state
$\state_i$. Finally, it is worth clarifying that in practise we do not
artificially limit the length of traces that the sequential planner
can consider. Moreover, in our setting there is always a finite
optimal trace.

@@

The DT session plans in a {\em small} abstract processes defined
according to the assumptive actions in the proceeding trace, and the
$(\pp{:init})$ declaration that models the underlying
belief-state. Targeted sensing is encouraged by augmenting the reward
model to reflect the value of knowing the truth about assumptions in
the trace proposed by the preceding sequential deliberation. In
detail, all rewards from the original POMDP are
retained. Additionally, for each $\assumptiveS{i}$ action scheduled by
the current trace, we have a {\em dual} $\assumptiveDT{i}$ so that
forall states $\state$:

@@

\noindent where $\$(T_i)$ (resp. $\hat\$(T_i)$) is a 
relatively\footnote{Compared to rewards from the original process
model.} small positive (negative) numeric quantity which captures the
utility the agent receives for correctly (incorrectly) rejecting an
assumption. Dual action execution is modelled as a self-transition --
I.e., the process remains in the state from which the dual is
executed.
%%
Furthermore, if \switchAction\ is the action that switched the system
to a DT session, then an assumption $\assumptiveS{i}$ is only {\em
active} if
\switchAction\ is not applicable in the sequential model unless
$\assumptiveS{i}$ is scheduled in the trace prefix
to \switchAction. We only include the duals of active assumptions in
DT sessions. For example, if the trace is:

@@

\noindent Taking the switching action \switchAction\ to be
$(\pp{look}~\pp{box}~\pp{office})$, we have that
$\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}))$ is not
active, and therefore exclude
$\actions^{\bullet}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}))$ from the
POMDP posed to the DT session.

>>>>

We only consider deterministic-action POMDPs, where all state
uncertainty is expressed in the $(\pp{:init})$ declaration. This
declaration is used by our approach to define the starting state for
sequential sessions, and the set of state-assumptions available to
sequential planning.  Writing \#\ if the value of a proposition is
unspecified, taking the $(\pp{:init})$ example from the previous
section, we have the following assumptions:

@@

\noindent Each assumption corresponds to one distinct {\em
relaxed} visitation of the root. Here, a conjunctive term is visited
iff its atomic subterms are visited, and zero or one of its immediate
probabilistic subterms are visited. For a sequential session, the
starting state corresponds to an assumption (usually unique) with
probability $1$. In our example, that starting state is:

@@

Encapsulating state-assumptions, we annotate the problem posed during
a sequential session with an assumptive action $\assumptiveS{i}$ for
each element $\prob_i (T_i)$, of each probabilistic term from
$(\pp{:init})$. Here, $\assumptiveS{i}$ can be executed if no
$\assumptiveS{j}$, $j \neq i$, has been executed from the same
probabilistic term, and, either
$(\pp{probabilistic}~..\prob_i~(T_i)..)$ is in the root conjunct, or
it occurs in $T_k$ for some executed $\assumptiveS{k}$.
%%
We add additional constraints that forbid the scheduling of
assumptions about facts after actions with preconditions, or effects,
that mention those facts. For example, the robot cannot assume it is
plugged into a power source immediately after it unplugs itself.
%%
Executing $\assumptiveS{i}$ in a state $\state$ effects a transition
to a successor state $\state^{T_i}$ with probability $\prob_i$, and
$\state^\bot$ with probability $1 - \prob_i$. Here, $\state^{T_i}$ is
the union of $\state$ with atomic terms from $T_i$. State
$\state^\bot$ is an added sink.

@@

\noindent  Here,  $\reward(\state_i, \action_i)$ is the instantaneous
reward ---i.e., can be negative for action costs, or positive for goal
achievement--- received for executing action $\action_i$ in state
$\state_i$. Finally, it is worth clarifying that in practise we do not
artificially limit the length of traces that the sequential planner
can consider. Moreover, for problems we consider there is always a
finite optimal trace.

@@

The DT session plans in a {\em small} abstract processes defined
according to the assumptive actions in the proceeding trace, and the
$(\pp{:init})$ declaration that models the underlying
belief-state. Targeted sensing is encouraged by augmenting the reward
model to reflect a heuristic value of knowing the truth about
assumptions in the trace. In detail, all rewards from the original
POMDP are retained. Additionally, for each $\assumptiveS{i}$ action
scheduled by the current trace, we have a {\em dual}
$\assumptiveDT{i}$ so that forall states $\state$:

@@

\noindent where $\$(T_i)$ (resp. $\hat\$(T_i)$) is a 
relatively\footnote{Compared to rewards from the original process
model.} small positive (negative) numeric quantity which captures the
utility the agent receives for correctly (incorrectly) rejecting an
assumption. A dual action can only be executed once, and corresponds
otherwise to a self-transition -- I.e., except in the applicability of
the dual, the process remains in the state from which a dual is
executed.
%%
Furthermore, if \switchAction\ is the action that switched the system
to a DT session, then an assumption $\assumptiveS{i}$ is only {\em
active} if
\switchAction\ is not applicable in the sequential model unless
$\assumptiveS{i}$ is scheduled in the trace prefix
to \switchAction. We only include the duals of active assumptions in
DT sessions. For example, if the trace is:

@@

\noindent Taking the switching action \switchAction\ to be
$(\pp{look}~\pp{box}~\pp{office})$, we have that
$\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}))$ is not
active, and therefore exclude it from the
POMDP posed to the DT session.
